Put Title here

### 2018-10-30 General WLD combinatorics.tex
added a new section 3.1 (containing Sian's lemma) and a few small results to section 3.2 (which used to be 3.1)

\section{Geometry of Wilson Loop diagram}\label{sec GN algorithm}
Since Wilson loop diagrams correspond to positroids, it is natural to understand the subspace of $\Gr(|\cP|, n)$ they define.

@@

@@

If the propagator $p$ is removed at vertex $i$ according to the algorithm for $I_a$, we say that $p$ \emph{contributes} $i$ to $I_a$.

@@

@@

\subsection{Some definitions and results that should probably be moved to section 3 but I don't know if anyone's editing that right now}

It will often be useful to refer to the vertex label contributed by a propagator to a Grassmann necklace term. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, n)$ be an admissible Wilson loop diagram. For each $i \in \interval{1}{n}$, define a function $I_i : \cP \longrightarrow \interval{1}{n}$ by
\[I_i(p) := \text{the vertex label that $p$ contributes to $I_i$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$.
\end{dfn}
\todo{Depending where this ends up, possibly include an example here.}

\begin{dfn}\label{props inside p}
Let $p$ be a propagator supported on $(i,j)$. The region bounded by $p$ and the edge of the Wilson loop diagram from $i$ to $j+1$ is refered to as the region {\bf inside} $p$, and the complement is the region {\bf outside }$p$. The set of propagators inside (resp. outside) $p$ are those lying in the region inside (resp. outside) $p$, excluding $p$ itself.
\end{dfn}
Note that this depends on the order of $i$ and $j$: if we view $p$ as supported on $(j,i)$ then the regions inside and outside $p$ are swapped.

\subsection{actual beginning of section 4; this subsection header can go away once the results above have found their place}

>>>>

\section{Geometry of Wilson Loop diagram}\label{sec GN algorithm}
Since Wilson loop diagrams correspond to positroids, it is natural to study the subspace of $\Gr(|\cP|, n)$ they define.

@@

Throughout this section, $W = (\cP,n)$ is an admissible Wilson loop diagram with $k$ propagators.

\subsection{Some setup results about propagators}

\note{rank probably isn't a good word; do we have a better one?}
\begin{dfn}\label{vertex rank}
Let $i$ be a vertex in $W$. We define the {\bf rank} of $i$ to be $|Prop(i)|$, i.e. the number of propagators supported on $i$.
\end{dfn}

\begin{dfn}\label{props inside p}
Let $p$ be a propagator supported on $(i,j)$ in $W$. The region bounded by $p$ and the edge of the Wilson loop diagram from $i$ to $j+1$ is refered to as the region {\bf inside} $p$, and the complement is the region {\bf outside }$p$. The set of propagators inside (resp. outside) $p$ are those lying in the region inside (resp. outside) $p$, excluding $p$ itself.
\end{dfn}
Note that this depends on the order of $i$ and $j$: if we view $p$ as supported on $(j,i)$ (for example, if we are working with the $<_j$ order on vertices) then the regions inside and outside $p$ are swapped.

\begin{dfn}\note{right now this doesn't depend on the order of $i$ and $j$, but that might matter later on?}
Let $p$ be a propagator supported on the vertices $(i,i+1,j,j+1)$.  Define the \textit{length} of $p$ to be 
\[\ell(p) = min\big\{|[i+1,j]|,|[j+1,i]|\big\}.\]
\end{dfn}

\begin{rmk}\label{rem:props of length 2 and 3}
The following observations about propagators of short length in $W$ are easily verified:
\begin{itemize}
\item If every vertex in $W$ supports at least one propagator, then $W$ admits at least one propagator of length 2.
\item If $p = (i,i+1,i+3,i+4)$ is a propagator of length 3, then the middle vertex $i+2$ supports at most one propagator.
\end{itemize}
\end{rmk}

\begin{dfn}
Define the restriction of $W$ to a region $[i,j]$ to be the subdiagram obtained by forgetting all propagators not wholly supported on vertices in the cyclic interval $[i,j]$.  Denote this diagram by $W_{[i,j]}$.
\end{dfn}

If $p$ is a propagator supported on $(i,i+1,j,j+1)$, then we can consider two obvious restrictions of $D$, namely $D_{[i,j+1]}$ and $D_{[j,i+1]}$.  Clearly all other propagators $q\neq p$ appear in exactly one of these subdiagrams, while $p$ appears in both.  In this case, we will say that $p$ \textit{bounds the regions} $[i,j+1]$ and $[j,i+1]$; the ordering of the start and end points allow us to identify one or the other region unambiguously.

The following lemma establishes certain configurations of propagators that must exist in any admissible diagram with no unsupported vertices. We make use of this result in several induction proofs below.

\begin{lem}\label{lem sian}
  Let $W$ be an admissible WLD in which each vertex supports at least one propagator.  Then exactly one of the following things occurs.
  \begin{enumerate}
    \item $W$ has a propagator of length 3, say $p =  (i,i+3)$, with a propagator of length 2 inside it.
    \item $W$ admits no propagators of length 3, and there exists a pair of propagators with the property that the first propagator is $(i, i+1, i+2, i+3)$, the second is $(i+2, i+3, i+4, i+5)$, and no other propagator ends on the edge $(i+2, i+3)$.
  \end{enumerate}
\end{lem}

\note{This is the proof of the old result pasted without modification so we can identify where it needs editing. For some reason the WLD was named $D$ instead of $W$.}
\begin{proof}
We will prove the following stronger result: that $D$ admits a pair of propagators of length 2 in the following configuration:
\begin{equation}\label{eq:2 prop config}
\begin{tikzpicture}[baseline=(current bounding box.east)]
\pgfmathsetmacro{\radius}{2}
\pgfmathsetmacro{\angle}{20}

\begin{scope}
\clip (-\radius*1.1,\radius*0.25) rectangle (\radius*1.1,\radius*1.25);

\begin{scope}
\clip (-\radius*1.1,\radius*0.45) rectangle (\radius*1.1,\radius*1.1);
\draw[-] (0,0) circle[radius=\radius cm];
\end{scope}

% vertices
\foreach \i in {2,3,...,7}{
	\draw (\angle*\i:\radius*0.98) --  (\angle*\i:\radius*1.02);
}

\begin{scope}
\clip (0,0) circle[radius=\radius];
\draw[propagator] (2.5*\angle:\radius*1.1 cm) to [bend left=90] (4.35*\angle:\radius*1.1 cm);
\draw[propagator] (4.65*\angle:\radius*1.1 cm) to [bend left=90] (6.5*\angle:\radius*1.1 cm);

\end{scope}

\node at (80:\radius*1.1) {\tiny $j$};
\node at (100:\radius*1.1) {\tiny $j+1$};
\node at (90:\radius*0.5) {\small $s_j = s_{j+1} = 2$};

\end{scope}
\end{tikzpicture}
\end{equation}

We will inductively construct a sequence of pairs of propagators $(p_r,q_r)$ satisfying: $\ell(p_r) = 2$, and $p_r$ either forms part of a configuration \eqref{eq:2 prop config} in $D$ or there is a propagator $q_r$ satisfying
\begin{itemize}
\item $\ell(q_r) \geq 4$.
\item $q_r$ bounds a region of $D$ which contains strictly fewer vertices than that bounded by $q_{r-1}$ and the restriction of $D$ to this region does not contain any of the propagators $p_1, \dots, p_r$.
\end{itemize}
Since $D$ contains finitely many vertices, this must eventually terminate in a configuration of the form \eqref{eq:2 prop config}.

Start by choosing a propagator $p_1$ of length 2 in $D$ (which exists by Remark~\ref{rem:props of length 2 and 3}), and write $(j_1,j_1+1,j_1+2,j_1+3)$ for its support.  If it is part of a configuration \eqref{eq:2 prop config} we are done, so suppose otherwise: since $s_{j_1} \geq 2$, there must exist a second propagator $q_1$ of length $\geq 4$ which starts on edge $(j_1,j_1+1)$.  This bounds a region $[i_1,j_1+1]$ which does not contain $p_1$.

Now suppose $q_{r}$ exists by the induction hypothesis and bounds a region $[i_r,j_r+1]$.  Consider the restricted diagram $D_r:=D_{[i_r,j_r+1]}$; clearly we must have $s_l \geq 2$ for all $l \in [i_r+2,j_r-1]$ (which is a non-empty interval since $\ell(q_r) \geq 4$), while $s_l\geq 1$ for $l \in \{i_r,i_r+1,j_r,j_r+1\}$.

By Remark~\ref{rem:props of length 2 and 3}, $D_r$ admits at least one propagator of length 2 and none of length 3.  Let $p_{r+1}$ be a propagator of length 2 in $D_r$; if it forms part of the configuration \eqref{eq:2 prop config} then we are done.  Otherwise, at least one endpoint of $p_{r+1}$ lies on an edge contained in the non-empty interval $[i_r+2,j_r-1]$, and so there is at least one other propagator connected to this edge.  Up to symmetry, there are three possibilities:

% this was 100% a productive use of my time

\[\begin{tabular}{ccc}
% possibility 1
\begin{tikzpicture}
\pgfmathsetmacro{\radius}{2}
\pgfmathsetmacro{\angle}{15}
\begin{scope}
\clip (-\radius*1.1,0) rectangle (\radius*1.1,\radius*1.1);

% vertices 
\foreach \i in {2,3,4,6,7,8,9,10}{
	\draw (\angle*\i:\radius*0.98) --  (\angle*\i:\radius*1.02);
}

% \dots 
\foreach \i in {1,2,3}{
	\node at (65 + \i*5:\radius cm) {$\cdot$};
}

\begin{scope} % this scope contains the propagators
\clip (0,0) circle[radius=\radius];

\draw[propagator] (\angle*2.4:\radius*1.2 cm) to [bend left=30] (\angle*9.7:\radius*1.2 cm); % q_r
\draw[propagator] (\angle*7.7:\radius*1.1 cm) to [bend left=90] (\angle*9.45:\radius*1.1 cm); % p_{r+1}

% grey prop begins here

\begin{scope}
\clip (\angle*7.5:\radius cm) circle[radius=\radius*0.45cm];
\draw[light-gray,propagator,seconddash] (\angle*6:\radius*1.02 cm) circle[x radius=\radius*0.4cm, y radius = \radius*0.3cm];
\end{scope}

\begin{scope}
\clip (0,0) circle[radius=\radius];
\clip (\angle*7.5:\radius cm) circle[radius=\radius*0.4cm];
\draw[light-gray,propagator,firstdash] (\angle*6:\radius*1.02 cm) circle[x radius=\radius*0.4cm, y radius = \radius*0.3cm];
\end{scope}

\begin{scope}
\clip (\angle*7.5:\radius cm) circle[radius=\radius*0.3cm];
\clip (0,0) circle[radius=\radius];
\draw[light-gray,propagator] (\angle*6:\radius*1.02 cm) circle[x radius=\radius*0.4cm, y radius = \radius*0.3cm];
\end{scope}
% end gray prop

\end{scope}

\node at (90:\radius*0.30 cm) {\tiny $q_r$};
\node at (\angle*8:\radius*0.73) {\tiny $p_{r+1}$};
\node[light-gray] at (\angle*4.9:\radius*0.77 cm) {\tiny $q_{r+1}$};
\node at (2*\angle:\radius*1.1) {\tiny $i_r$};
\node at (9*\angle:\radius*1.1) {\tiny $j_r$};

% circle/edges; needs to happen last to overlap gray
\begin{scope}
\clip (-\radius*1.1,0.9) rectangle (\radius*1.1,\radius*1.1);
\draw[-] (0,0) circle[radius=\radius cm];
\end{scope}

\end{scope} % ends the overall rectangle clipping
\end{tikzpicture}

&

% possibility 2
\begin{tikzpicture}
\pgfmathsetmacro{\radius}{2}
\pgfmathsetmacro{\angle}{15}
\begin{scope}
% overall clipping to make the size sensible
\clip (-\radius*1.1,0) rectangle (\radius*1.1,\radius*1.1);

% vertices
\foreach \i in {2,3,4,6,7,8,9,10}{
	\draw (\angle*\i:\radius*0.98) --  (\angle*\i:\radius*1.02);
}

% \dots
\foreach \i in {1,2,3}{
	\node at (65 + \i*5:\radius cm) {$\cdot$};
}

\begin{scope} % props
\clip (0,0) circle[radius=\radius cm]; 
\draw[propagator] (\angle*2.4:\radius*1.2 cm) to [bend left=30] (\angle*9.6:\radius*1.2 cm); % q_r
\draw[propagator] (\angle*6.7:\radius*1.1 cm) to [bend left=90] (\angle*8.6:\radius*1.1 cm); % p_{r+1}

%% grey prop starts here

\clip (\angle*6.5:\radius cm) circle[radius=\radius*0.45cm];
\draw[light-gray,propagator,seconddash] (\angle*4.5:\radius*1.3 cm) circle[y radius = 0.5*\radius cm,x radius=\radius*0.7cm];
\end{scope}

\begin{scope}
\clip (0,0) circle[radius=\radius];
\clip (\angle*6.5:\radius cm) circle[radius=\radius*0.4cm];
\draw[light-gray,propagator,firstdash] (\angle*4.5:\radius*1.3 cm) circle[y radius = 0.5*\radius cm,x radius=\radius*0.7cm];
\end{scope}

\begin{scope}
\clip (\angle*6.5:\radius cm) circle[radius=\radius*0.3cm];
\clip (0,0) circle[radius=\radius];
\draw[light-gray,propagator] (\angle*4.5:\radius*1.3 cm) circle[y radius = 0.5*\radius cm,x radius=\radius*0.7cm];
\end{scope}
% grey prop ends here

\node at (90:\radius*0.32 cm) {\tiny $q_r$};
\node at (\angle*7.75:\radius*0.75) {\tiny $p_{r+1}$};
\node[light-gray] at (\angle*4.5:\radius*0.7 cm) {\tiny $q_{r+1}$};
\node at (2*\angle:\radius*1.1) {\tiny $i_r$};
\node at (9*\angle:\radius*1.1) {\tiny $j_r$};

% circle/edges
\begin{scope}
\clip (-\radius*1.1,0.9) rectangle (\radius*1.1,\radius*1.1);
\draw[-] (0,0) circle[radius=\radius cm];
\end{scope}

\end{scope} % ends overall clip
\end{tikzpicture}

&

% possiblitiy 3
\begin{tikzpicture}
\pgfmathsetmacro{\radius}{2}
\pgfmathsetmacro{\angle}{12}
\begin{scope}
\clip (-\radius*1.1,0) rectangle (\radius*1.1,\radius*1.1);

% vertices
\foreach \i in {2,3,6,7,8,9,12,13}{
	\draw (\angle*\i:\radius*0.98) --  (\angle*\i:\radius*1.02);
}

% \dots
\foreach \i in {1,2,3}{
	\node at (44 + \i*5:\radius cm) {$\cdot$};
	\node at (116 + \i*5:\radius cm) {$\cdot$};
}

\begin{scope} % props
\clip (0,0) circle[radius=\radius cm];
\draw[propagator] (\angle*2.4:\radius*1.2 cm) to [bend left=30] (\angle*12.6:\radius*1.2 cm); % q_r
\draw[propagator] (\angle*6.5:\radius*1.1 cm) to [bend left=100] (\angle*8.5:\radius*1.05 cm); % p_{r+1}

%  grey prop 1
% not q_{r+1}
\begin{scope}
\clip (\angle*6.5:\radius cm) circle[radius=\radius*0.45cm];
\draw[light-gray,propagator,seconddash] (\angle*6.5:\radius*1.1 cm) to [bend left=90] (\angle*12.5:\radius*1.1cm);
\end{scope}

\begin{scope}
\clip (\angle*6.5:\radius cm) circle[radius=\radius*0.4cm];
\draw[light-gray,propagator,firstdash] (\angle*6.5:\radius*1.1 cm) to [bend left=90] (\angle*12.5:\radius*1.1cm);
\end{scope}
\begin{scope}
\clip (\angle*6.5:\radius cm) circle[radius=\radius*0.3cm];
\draw[light-gray,propagator] (\angle*6.5:\radius*1.1 cm) to [bend left=90] (\angle*12.5:\radius*1.1cm);
\end{scope}
% grey prop 1 ends here

% grey prop 2
% this is q_{r+1}
\begin{scope}
\clip (\angle*8.5:\radius cm) circle[radius=\radius*0.45cm];
\draw[propagator,seconddash,light-gray] (\angle*8.8:\radius*1.1 cm) to [bend left=90] (\angle*12:\radius*1.1cm);
\end{scope}

\begin{scope}
\clip (\angle*8.5:\radius cm) circle[radius=\radius*0.4cm];
\draw[propagator,firstdash,light-gray] (\angle*8.8:\radius*1.1 cm) to [bend left=90] (\angle*12:\radius*1.1cm);
\end{scope}

\begin{scope}
\clip (\angle*8.5:\radius cm) circle[radius=\radius*0.3cm];
\draw[light-gray,propagator] (\angle*8.8:\radius*1.1 cm) to [bend left=90] (\angle*12:\radius*1.1cm);
\end{scope}

% grey prop 2 ends here

\end{scope}

\node at (90:\radius*0.15 cm) {\tiny $q_r$};
\node at (\angle*7.5:\radius*0.83) {\tiny $p_{r+1}$};
\node[light-gray] at (\angle*11:\radius*0.75 cm) {\tiny $q_{r+1}$};

\node at (2*\angle:\radius*1.1) {\tiny $i_r$};
\node at (12*\angle:\radius*1.1) {\tiny $j_r$};

% overall circle
\begin{scope}
\clip (-\radius*1.1,0.7) rectangle (\radius*1.1,\radius*1.1);
\draw[-] (0,0) circle[radius=\radius cm];
\end{scope}

\end{scope} % ends overall clip
\end{tikzpicture}

\end{tabular}
\]
In each case, the diagram indicates how to choose $q_{r+1}$ satisfying the induction hypothesis.  In the third case, note that \textit{both} ends of $p_{r+1}$ must share their edge with a propagator of length $\geq 4$; the non-crossing hypothesis on propagators guarantees that at least one will bound a region not containing $p_{r+1}$.
\end{proof}

@@

If the propagator $p$ is removed at vertex $i$ according to the algorithm for $I_a$, we say that $p$ \emph{contributes} $i$ to $I_a$. Notationally, we represent this by allowing the $I_i$ symbol to represent a function as well as a set, as follows:
\begin{dfn}\label{def I_i as a function}
Let $W = (\cP, n)$ be an admissible Wilson loop diagram. For each $a \in [n]$, define a function $I_a : \cP \longrightarrow [n]$ by
\[I_a(p) := \text{the vertex label that $p$ contributes to $I_a$ in Algorithm~\ref{alg:put GN on WLD},}\]
for each $p \in \cP$.
\end{dfn}

@@

In an arbitrary Grassmann necklace, it is possible for an index $i$ to appear in no terms of the Grassmann necklace (a {\em loop}) or in all terms of the necklace (a {\em coloop}). Clearly any vertex of $W$ that supports no propagators will yield a loop [ref][check: is this iff?], but with Theorem [ref][cyclic intervals thm] in hand we can now easily verify that admissible Wilson loop diagrams do not admit coloops.

@@
